Mendelian randomization for causal inference in Genome-wide association studies

### 2022-02-12 mr-review.Rmd
echo = FALSE
added more about mendel

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

@@

To understand the origin of the term "Mendelian randomization", it is helpful to 
recall Mendel's inheritance laws. Under the assumption that alleles segregate 
randomly from parent to offspring, the offspring genotypes are unlikely to be 
associated with confounders of the exposure-outcome relationship. Additionally,
reverse causation is ruled out by the observation that germ-line genotypes are 
fixed at conception and, thus, precede other observed variables.

@@

# Binary, Count, and Time-to-event outcomes

@@

>>>>

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(magrittr)
```

@@

To understand the origin of the term "Mendelian randomization", it is helpful to 
recall Mendel's inheritance laws [@mendel1865versuche]. Under the assumption that alleles segregate 
randomly from parent to offspring, the offspring genotypes are unlikely to be 
associated with confounders of the exposure-outcome relationship. Additionally,
reverse causation is unanticipated since germ-line genotypes are 
fixed at conception and, thus, precede realization of other observed variables. Reverse causation 
occurs when the outcome at an early time point has a causal effect on the exposure variable. While 
reverse causation is often assumed to not occur in MR study designs, @burgess_are_2021 points out 
several scenarios where this assumption may not hold [@davey_smith_mendelian_2003].

@@

# Binary, Count, and Time-to-event outcomes and exposures

@@

While we've focused on continuous outcomes and continuous exposures, some
data are more naturally treated as binary, count, or time-to-event variables. MR methods 
for non-continuous outcomes and exposures is an active area of research. @burgess_improving_2012 and @allman_novel_2021 
developed MR strategies for binary exposures and binary outcomes. @burgess_improving_2012 extended 
methods from @wald_fitting_1940 by fitting two logistic regressions, one for the outcome and one for 
the exposure. They then calculated the causal effect estimator as the ratio of the two logistic 
regression coefficient estimators. @allman_novel_2021 devised a new MR method for binary outcome and 
binary exposure by drawing on @gauvreau_analysis_1997 and treating the exposure and outcome as correlated binary variables. Writing the likelihood for this model, they implemented an iterative optimization algorithm to simultaneously infer the causal effect parameter and other parameters.
